Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-07 Manuscript/TEM_draft.tex
Edits BC explanation

\subsubsection{Summarizing the comparisons}
Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree's topology in a Total Evidence framework. We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution).

%Introducing Bhattacharyya Coefficient
To evaluate these effects we measured similarities among different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for all the analyses above using the Bhattacharyya Coefficient \citep{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).

%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
For assessing the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the Normalised Robinson-Foulds and Normalised Triplets distance distributions, for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%). For example, $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc. This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells). We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}). We repeated this procedure for all the "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "best" tree's topology, we compared the similarity between the methods. We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table~\ref{Tab_Results-Difference_methods}. When two methods perform similarly in their ability to recover the "correct" tree topology (i.e. they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.

%Questionish §
We use the Bhattacharyya Coefficient in a similar way as a two-sided t-test to detect significant differences between the distributions of the normalised distance between the "best" tree and the "missing data" trees (for both metrics).

>>>>

\subsection{Testing the effects of the missing data parameters on topological recovery} % NC: I think this is really a separate section to tree comparison.
Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree topology in a Total Evidence framework. We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution).

We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details). Note that this is  equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means. To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc. (see Figure % FIGURE
for more details). This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells). We performed this procedure separately for each tree inference and uncertainty method. When two combinations of missing data parameters have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.

% NC: Maybe put this in the results?
% We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}).

To assess the effect of the different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the "best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods i.e. Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc. (see Figure % FIGURE
for more details). Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons. When two tree inference or uncertainty methods have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
